# Statistical Rethinking Chapter 4 problems

__Name:__


# For 03/17/16

## 4E1 In the model definition below, which line is the likelihood?


likelihood = binomal(number of times seen, number of samples)
likelihood = yi ∼ Normal(μ, σ)
The value yi is distrbuted normally by the number of samples and the probality which is givin by the mean and variance....

σ prior = σ ∼ Uniform(0, 10) uniform prior between zero and 10
μ prior = μ ∼ Normal(0, 10) gaussian prior centered on 0 and
prios are problistic not known with certinity


## 4E2 In the model definition just above, how many parameters are in the posterior distribution? 

10



## 4E3 Using the model definition above, write down the appropriate form of Bayes’ theorem that includes the proper likelihood and priors.

pr(μ,σ|h) = ∏ Normal(hi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10) / ∫ ∫ ∏ Normal(hi|μ, σ)Normal(μ|0, 10)Uniform(σ|0, 50)dμdσ

## 4M1 For the model definition below, simulate observed heights from the prior (not the posterior).

yi ∼ Normal(μ, σ)
μ ∼ Normal(0, 10) 
σ ∼ Uniform(0, 10)


```{r}

flist <- alist(
    height ~ dnorm( mu , sigma ) ,
    mu ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
)
sample_mu <- rnorm( 1e4 , 0 , 10 )
sample_sigma <- runif( 1e4 , 0 , 10 )

prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )

yi ∼ Normal(μ, σ)
μ ∼ Normal(0, 10) σ ∼ Uniform(0, 10)
```

## 4M2 Translate the model just above into a map formula.

```{r}
library(rethinking)
data(Howell1)
d <- Howell1

m4.2 <- map(
        alist(
            height ~ dnorm( mu , sigma ) ,
            mu ~ dnorm( 0 , 10 ) ,
            sigma ~ dunif( 0 , 10 )
),
        data=d )
precis( m4.2 )




```






# For 03/24/16

## 4E4 In the model definition below, which line is the linear model? 

This is the linear model : μi =α+βxi

yi ∼ Normal(μ, σ)
μi =α+βxi
α ∼ Normal(0, 10) 
β ∼ Normal(0, 1)
σ ∼ Uniform(0, 10)



## 4E5 In the model definition just above, how many parameters are in the posterior distribution?
 3 - μ, σ and μi? 

## 4M3 Translate the map model formula below into a mathematical model definition.

```{r}
flist <- alist(
    y ~ dnorm( mu , sigma ),
    mu <- a + b*x,
    a ~ dnorm( 0 , 50 ),
    b ~ dunif( 0 , 10 ),
    sigma ~ dunif( 0 , 50 )
)
```

yi ∼ Normal(μ, σ)
μi =α+βxi
α ∼ Normal(0, 50) 
β ∼ Normal(0, 10)
σ ∼ Uniform(0, 50)


## 4M4 A sample of students is measured for height each year for 3 years.After the third year,you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.

yi ∼ Normal(μ, σ)
* linear model
μi =α+βyr
* yaxis
α ∼ Normal(63, 100) 
* slope
β ∼ Normal(0, 10)
* variance
σ ∼ Uniform(0, 50)


 I’ve widened the prior for α, since as you’ll see it is common for the intercept in a linear model to swing a long way from the mean of the outcome variable. The flat prior here with a huge standard deviation will allow it to move wherever it needs to. 

β = 0, weight has no relationship to height. So many people see it as conservative assumption. And such a prior will pull probability mass towards zero, leading to more conservative estimates

50 is a pretty broud deviance 

## 4M5 now suppose I tell you that the average height in the first year was 120 cm and that every student got taller each year. Does this information lead you to change your choice of priors? How?

yi ∼ Normal(μ, σ)
* intercept
μi =α+βyr
* yaxis 
α ∼ Normal(120, 50) 
* slope
β ∼ Normal(0, 10)
* variance
σ ∼ Uniform(0, 50)



## 4M6 Now suppose I tell you that the variance among heights for students of the same age is never more than 64cm. How does this lead you to revise your priors?


yi ∼ Normal(μ, σ)
* intercept
μi =α+βyr
* yaxis 
α ∼ Normal(120, 50) 
* slope
β ∼ Normal(0, 10)
* variance
σ ∼ Uniform(0, 64)






# For 03/31/16

## 4H1

## 4H2

## 4H3
