# Statistical Rethinking Chapter 4 problems

__Name:__


# For 03/17/16

## 4E1 In the model definition below, which line is the likelihood?


likelihood = binomal(number of times seen, number of samples)
likelihood = yi ∼ Normal(μ, σ)
The value yi is distrbuted normally by the number of samples and the probality which is givin by the mean and variance....

σ prior = σ ∼ Uniform(0, 10) uniform prior between zero and 10
μ prior = μ ∼ Normal(0, 10) gaussian prior centered on 0 and
prios are problistic not known with certinity


## 4E2 In the model definition just above, how many parameters are in the posterior distribution? 

10



## 4E3 Using the model definition above, write down the appropriate form of Bayes’ theorem that includes the proper likelihood and priors.

pr(μ,σ|h) = ∏ Normal(hi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10) / ∫ ∫ ∏ Normal(hi|μ, σ)Normal(μ|0, 10)Uniform(σ|0, 50)dμdσ

## 4M1 For the model definition below, simulate observed heights from the prior (not the posterior).

yi ∼ Normal(μ, σ)
μ ∼ Normal(0, 10) 
σ ∼ Uniform(0, 10)


```{r}

flist <- alist(
    height ~ dnorm( mu , sigma ) ,
    mu ~ dnorm( 0 , 10 ) ,
    sigma ~ dunif( 0 , 10 )
)
sample_mu <- rnorm( 1e4 , 0 , 10 )
sample_sigma <- runif( 1e4 , 0 , 10 )

prior_h <- rnorm( 1e4 , sample_mu , sample_sigma )
dens( prior_h )

yi ∼ Normal(μ, σ)
μ ∼ Normal(0, 10) σ ∼ Uniform(0, 10)
```

## 4M2 Translate the model just above into a map formula.

```{r}
library(rethinking)
data(Howell1)
d <- Howell1

m4.2 <- map(
        alist(
            height ~ dnorm( mu , sigma ) ,
            mu ~ dnorm( 0 , 10 ) ,
            sigma ~ dunif( 0 , 10 )
),
        data=d )
precis( m4.2 )




```






# For 03/24/16

## 4E4

## 4E5

## 4M3

## 4M4

## 4M5

## 4M6

# For 03/31/16

## 4H1

## 4H2

## 4H3
