Chapter 6 notes
========================================================

## Section 1-2

```{r}
sppnames <- c( "afarensis","africanus","habilis","boisei",
    "rudolfensis","ergaster","sapiens")
brainvolcc <- c( 438 , 452 , 612, 521, 752, 871, 1350 )
masskg <- c( 37.0 , 35.5 , 34.5 , 41.5 , 55.5 , 61.0 , 53.5 )
d <- data.frame( species=sppnames , brain=brainvolcc , mass=masskg )

m6.1 <- lm( brain ~ mass , data=d )
summary(m6.1)
##r^2
1 - var(resid(m6.1))/var(d$brain)



m6.2 <- lm( brain ~ mass + I(mass^2) , data=d )
m6.3 <- lm( brain ~ mass + I(mass^2) + I(mass^3) , data=d )
m6.4 <- lm( brain ~ mass + I(mass^2) + I(mass^3) + I(mass^4) ,
    data=d )
m6.5 <- lm( brain ~ mass + I(mass^2) + I(mass^3) + I(mass^4) +
    I(mass^5) , data=d )
m6.6 <- lm( brain ~ mass + I(mass^2) + I(mass^3) + I(mass^4) +
    I(mass^5) + I(mass^6) , data=d )



# fit model with lm
m6.1 <- lm( brain ~ mass , d )
# compute deviance by cheating
(-2) * logLik(m6.1)

library(rethinking)

# standardize the mass before fitting
d$mass.s <- (d$mass-mean(d$mass))/sd(d$mass)
m6.8 <- map(
    alist(
        brain ~ dnorm( mu , sigma ) ,
        mu <- a + b*mass.s
),
data=d , start=list(a=mean(d$brain),b=0,sigma=sd(d$brain)) ,
        method="Nelder-Mead" )
# extract MAP estimates
theta <- coef(m6.8)
# compute deviance
dev <- (-2)*sum( dnorm(
  d$brain ,
mean=theta[1]+theta[2]*d$mass.s ,
sd=theta[3] ,
log=TRUE ) )
dev

```

```{r}

d1$rround(d1$age)

m <- map(
    alist(
        dist ~ dnorm(mu,sigma),
        mu <- a + b*speed,
        a ~ dnorm(0,100),
        b ~ dnorm(0,10),
        sigma ~ dunif(0,30)
    ) , data=cars )

post <- extract.samples(m,n=1000)


### use the alpha at sample 1 from the extracted sample from postioior dist and   beta value at sample one * all the car speeds the dataset for each post dist sample 1-1000
### the second line calculated the loglikehood for each valule from the posteior as seen in WAIC
n_samples <- 1000
ll <- sapply( 1:n_samples ,
    
              function(s) {
        mu <- post$a[s] + post$b[s]*cars$speed
        dnorm( cars$dist , mu , post$sigma[s] , log=TRUE )
})


