---
title: "Chp_14 homework"
output:
  html_document:
    keep_md: true
---


# 14E1. Rewrite the Oceanic tools model (from Chapter 10) below so that it assumes measured error on the log population sizes of each society.

Dest,i ∼ Poisson(μi)
logμi =α+βlogPi

Dobs,i ∼ Normal(Dest,i, Dse,i)


α ∼ Normal(0, 10) 
β ∼ Normal(0, 1)


## 14E2. Rewrite the same model so that it allows imputation of missing values for log population.  Tere aren’t any missing values in the variable, but you can still write down a model formula that would imply imputation, if any values were missing.

Dest,i ∼ Poisson(μi)
logμi =α+βlogPi
Pi ∼ Normal(ν, σN)
Dobs,i ∼ Normal(Dest,i, Dse,i)


α ∼ Normal(0, 10) 
β ∼ Normal(0, 1)


## 14M1. Using the mathematical form of the imputation model in the chapter, explain what is being assumed about how the missing values were generated.

Both models assume that the data is not missing due to bais in collection.
Both models are assuming that the missing data is normaly distributed. 
The second model assumes that their is a linear relationship.

## 14M2. In earlier chapters, we threw away cases from the primate milk data, so we could use the neocortex variable. Now repeat the WAIC model comparison example from Chapter 6, but use im- putation on the neocortex variable so that you can include all of the cases in the original data.  e simplest form of imputation is acceptable. How are the model comparison results a effcted by being able to include all of the cases?

```{r}
library(rethinking)
data(milk)
d <- milk
d$neocortex.prop <- d$neocortex.perc / 100
d$logmass <- log(d$mass)

# prep data
data_list <- list(
    kcal = d$kcal.per.g,
    neocortex = d$neocortex.prop,
    logmass = d$logmass )
# fit model
m14.3 <- map2stan(
    alist(
        kcal ~ dnorm(mu,sigma),
        mu <- a + bN*neocortex + bM*logmass,
        neocortex ~ dnorm(nu,sigma_N),
        a ~ dnorm(0,100),
        c(bN,bM) ~ dnorm(0,10),
        nu ~ dnorm(0.5,1),
        sigma_N ~ dcauchy(0,1),
        sigma ~ dcauchy(0,1)
),
data=data_list , iter=1e4 , chains=2 )



dcc <- d[ complete.cases(d$neocortex.prop) , ]
data_list_cc <- list(
    kcal = dcc$kcal.per.g,
    neocortex = dcc$neocortex.prop,
    logmass = dcc$logmass )
# fit model
m14.3cc <- map2stan(
    alist(
        kcal ~ dnorm(mu,sigma),
        mu <- a + bN*neocortex + bM*logmass,
        a ~ dnorm(0,100),
        c(bN,bM) ~ dnorm(0,10),
        sigma ~ dcauchy(0,1)
),
    data=data_list_cc , iter=1e4 , chains=2 )
precis(m14.3cc)


compare(m14.3cc,m14.3)

```

# The model without missing data is favoried but has a larger SE as expected

## 14M3. Repeat the divorce data measurement error models, but this time double the standard errors. Can you explain how doubling the standard errors impacts inference?


```{r}
library(rethinking)
data(WaffleDivorce)
d <- WaffleDivorce


dlist <- list(
    div_obs=d$Divorce,
    div_sd=d$Divorce.SE,
    R=d$Marriage,
    A=d$MedianAgeMarriage
)


m14.1 <- map2stan(
    alist(
        div_est ~ dnorm(mu,sigma),
        mu <- a + bA*A + bR*R,
        div_obs ~ dnorm(div_est,div_sd),
        a ~ dnorm(0,10),
        bA ~ dnorm(0,10),
        bR ~ dnorm(0,10),
        sigma ~ dcauchy(0,2.5)
),
data=dlist ,
start=list(div_est=dlist$div_obs) ,
WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
control=list(adapt_delta=0.95) )

```

# double standerd error

```{r}
dlist2SE <- list(
    div_obs=d$Divorce,
    div_sd=d$Divorce.SE * 2,
    R=d$Marriage,
    A=d$MedianAgeMarriage
)


m14.M3 <- map2stan(
    alist(
        div_est ~ dnorm(mu,sigma),
        mu <- a + bA*A + bR*R,
        div_obs ~ dnorm(div_est,div_sd),
        a ~ dnorm(0,10),
        bA ~ dnorm(0,10),
        bR ~ dnorm(0,10),
        sigma ~ dcauchy(0,2.5)
),
data=dlist2SE ,
start=list(div_est=dlist2SE$div_obs) ,
WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
control=list(adapt_delta=0.95) )


```


## 14H1. The data in data(elephants) are counts of matings observed for bull elephants of differing ages.  There is a strong positive relationship between age and matings. However, age is not always assessed accurately. First,  t a Poisson model predicting MATINGS with AGE as a predictor. 


```{r}
data(elephants)
e <- elephants
summary(e)

m14.H1a <- map2stan(
    alist(
        MATINGS ~ dpois( lambda ),
        log(lambda) <- a + bA*AGE,
        a ~ dnorm(0,50),
        bA ~ dnorm(0,5)
), data=e )

precis(m14.H1a)

```

## Second, assume that the observed AGE values are uncertain and have a standard error of ±5 years. Re-estimate the relationship between MATINGS and AGE, incorporating this measurement error. Compare the inferences of the two models.


```{r}
elist <- list(
    MATINGS=e$MATINGS,
    age_sd=rep(5,length(e$AGE)),
    age_obs=e$AGE
)

m14.H1b <- map2stan(
    alist(
        MATINGS ~ dpois( lambda ),
        log(lambda) <- a + bA*age_est[i],
        age_obs ~ dnorm(age_est,age_sd),       
        a ~ dnorm(0,50),
        bA ~ dnorm(0,5)
), data=elist,
start=list(age_est=elist$age_obs) ,
WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
control=list(adapt_delta=0.95) )


precis(m14.H1b, depth=2)
WAIC(m14.H1a,m14.H1b)

```


## 14H2. Repeat the model  fitting problem above, now increasing the assumed standard error on AGE. How large does the standard error have to get before the posterior mean for the coeffcient on AGE reaches zero?

```{r}

elist2 <- list(
    MATINGS=e$MATINGS,
    age_sd=rep(10,length(e$AGE)),
    age_obs=e$AGE
)

m14.H2 <- map2stan(
    alist(
        MATINGS ~ dpois( lambda ),
        log(lambda) <- a + bA*age_est[i],
        age_obs ~ dnorm(age_est,age_sd),       
        a ~ dnorm(0,50),
        bA ~ dnorm(0,5)
), data=elist2,
start=list(age_est=elist2$age_obs) ,
WAIC=FALSE , iter=5000 , warmup=1000 , chains=2 , cores=2 ,
control=list(adapt_delta=0.95) )

precis(m14.H2)
```


