---
title: "Chp_13 homework"
output:
  html_document:
    keep_md: true
---

Probblems:


## 13E1. Add to the following model varying slopes on the predictor x. 

yi ∼ Normal(μi, σ)
μi = αgroup[i] + βx[i] 

[αgroup, βx] ∼ MVNormal([α
                         β ], S)
                         
S = (σα, 0 ` R(σα, 0
0, σβ)   0, σβ)

α ∼ Normal(0, 10)
β ∼ Normal(0, 1)
σ ∼ HalfCauchy(0, 2)
σα ∼ HalfCauchy(0, 2)
R ~ LKJcorr(2)

## 13E2.  Think up a context in which varying intercepts will be positively correlated with varying slopes.
## Provide a mechanistic explanation for the correlation.

### cafe example has a negative correlation between intercept and slope
cafe- average morrning wait 5, 20
slope - differnce between afternoon and morning wait higher, lower diff cause always busy

### plant height and developmental stage at measurment
intercept - average height of plants or growth tall <--- short
slope - difference between height measurments for young or older plants (measuring at recording diff developmental stages)

The shorter plants will remain with smaller growth and a samller difference
The larger plants will have more growth inbetween these stages and a larger difference


## 13E3. When is it possible for a varying slopes model to have fewerer effective parameters(as estimated by WAIC or DIC) than the corresponding model with  fixed (unpooled) slopes? Explain.

There are fewer effective paramters when you are using uncentered.
Does this have to do with rather than sampling from a distribution it is built into the modeling.. making the overall number of test/iterations need to run better. Its regulazing the information more towards what the data says? 


## 13M1
## Repeat the café robot simulation from the beginning of the chapter. This time,set rho to zero, so that there is no correlation between intercepts and slopes. How does the posterior distribution of the correlation reflect this change in the underlying simulation?

## simulation with rho set to zero, so that there is no correlation between intercepts and slopes.


```{r}
library(rethinking)
## simulate averages from populations / create fake dataset
a <- 3.5
b <- (-1)
sigma_a <- 1
sigma_b <- 0.5
rho <- (0)
N_cafes <- 20

## make two dimensional multivatiate gaussian
## average info for intercept and slope
Mu <- c( a , b )
### sd for intercept and slope 
cov_ab <- sigma_a*sigma_b*rho
Sigma <- matrix( c(sigma_a^2,cov_ab,cov_ab,sigma_b^2) , ncol=2)


### select from multivariate gaussian
library(MASS)
### select from multivariate gaussian
set.seed(5) # used to replicate example
vary_effects <- mvrnorm( N_cafes , Mu , Sigma )

a_cafe <- vary_effects[,1]
b_cafe <- vary_effects[,2]


N_visits <- 10
afternoon <- rep(0:1,N_visits*N_cafes/2)
cafe_id <- rep( 1:N_cafes , each=N_visits )




mu <- a_cafe[cafe_id] + b_cafe[cafe_id]*afternoon
sigma <- 0.5  # std dev within cafes
wait <- rnorm( N_visits*N_cafes , mu , sigma )
d2 <- data.frame( cafe=cafe_id , afternoon=afternoon , wait=wait )




m13.m1 <- map2stan(
    alist(
        wait ~ dnorm( mu , sigma ),
        mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
        c(a_cafe,b_cafe)[cafe] ~ dmvnorm2(c(a,b),sigma_cafe,Rho),
        a ~ dnorm(0,10),
        b ~ dnorm(0,10),
        sigma_cafe ~ dcauchy(0,2),
        sigma ~ dcauchy(0,2),
        Rho ~ dlkjcorr(2)
),
data=d2 ,
iter=5000 , warmup=2000 , chains=2 )


```

### How does the posterior distribution of the correlation reflect this change in the underlying simulation?

```{r}
post <- extract.samples(m13.m1)
dens( post$Rho[,1,2] )
### these show a postive correlation rather than non between slope and intercept?

##intercept and slope
## morrining and afternoon wait time
## most traffic in
```

## 13M2. Fit this multilevel model to the simulated café data: 
# Wi ∼ Normal(μi, σ)
# μi = αcafé[i] + βcafé[i]Ai 
# αcafé ∼ Normal(α, σα)
# βcafé ∼Normal(β,σβ)
# α ∼ Normal(0, 10)
# β ∼ Normal(0, 10)
# σ ∼ HalfCauchy(0, 1)
# σα ∼ HalfCauchy(0, 1) σβ ∼HalfCauchy(0,1)
## Use WAIC to compare this model to the model from the chapter, the one that uses a multi-variate Gaussian prior. Explain the result.

```{r}

m13.m2 <- map2stan(
    alist(
        wait ~ dnorm( mu , sigma ),
        mu <- a_cafe[cafe] + b_cafe[cafe]*afternoon,
        a_cafe[cafe] ~ dnorm(a, sigma_alpha),
        b_cafe[cafe] ~ dnorm(b, sigma_beta),
        a ~ dnorm(0,10),
        b ~ dnorm(0,10),
        sigma ~ dcauchy(0,1),
        sigma_alpha ~ dcauchy(0,1),
        sigma_beta ~ dcauchy(0,1)
),
data=d2 ,
iter=5000 , warmup=2000 , chains=2 )

compare(m13.m1,m13.m2)

```

These models are pretty much the same, except I amd confused why the dcauchy vars differ...

## 13M3. Re-estimate the varying slopes model for the UCBadmit data, now using a non-centered pa- rameterization. Compare the eciency of the forms of the model, using n_eff. Which is better? Which chain sampled faster?

nonparamterized should have better model needing less parmaters and should sample faster we will test

```{r}
library(rethinking)
data(UCBadmit)
d <- UCBadmit
d$male <- ifelse( d$applicant.gender=="male" , 1 , 0 )
d$dept_id <- coerce_index( d$dept )

m13.3 <- map2stan(
    alist(
        admit ~ dbinom( applications , p ),
        logit(p) <- a_dept[dept_id] +
                    bm_dept[dept_id]*male,
        c(a_dept,bm_dept)[dept_id] ~ dmvnorm2( c(a,bm) , sigma_dept , Rho ),
        a ~ dnorm(0,10),
        bm ~ dnorm(0,1),
        sigma_dept ~ dcauchy(0,2),
        Rho ~ dlkjcorr(2)
),
data=d , warmup=1000 , iter=5000 , chains=4 , cores=3 )





m13.3NC <- map2stan(
    alist(
        admit ~ dbinom( applications , p ),
        logit(p) <- a_dept[dept_id] +
                    bm_dept[dept_id]*male,
        c(a_dept,bm_dept)[dept_id] ~ dmvnormNC(sigma_dept , Rho ),
        a ~ dnorm(0,10),
        bm ~ dnorm(0,1),
        sigma_dept ~ dcauchy(0,2),
        Rho ~ dlkjcorr(2)
),
data=d , warmup=1000 , iter=5000 , chains=4 , cores=3 )

#m13.3 There were 26 divergent iterations during sampling
#m13.3NC There were 4 divergent transitions after warmup
compare(m13.3, m13.3NC)

precis(m13.3, depth=2)


precis(m13.3NC, depth=2)
```


The data looked a little better, needed a litle less for N_eff, but WAIC showed two models were very similar. It did run much faster and have less divergernt chains

## 13H1. Let’s revisit the Bangladesh fertility data, data(bangladesh),from the practice problems for Chapter 12.

```{r}

data(bangladesh)
d <- bangladesh
head(d)

## clean up data
d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))
## Now there are 60 values, contiguous integers 1 to 60.

colnames(d) <- gsub("\\.", "_", colnames(d))
d$number_of_women_in_distict <- rep(1,1934)

head(d)

```
## Fit a model with both varying intercepts by district_id and varying slopes of urban by district_id. You are still predicting use.contraception. 

```{r}

m13.h1 <- map2stan(
    alist(
        use_contraception ~ dbinom( 1 , p ) ,
        logit(p) <- a_district[district_id] + b_district[district_id]*urban,
        c(a_district,b_district)[district_id] ~   dmvnorm2(c(a,b),sigma_dist,Rho),
        Rho ~ dlkjcorr(2),
        c(a,b) ~ dnorm(0,1) ,
        c(sigma,sigma_dist) ~ dcauchy(0,1)
    ), data=d , iter=4000 , chains=4)

precis(m13.h1, depth=2)
```
## Inspect the correlation between the intercepts and slopes. Can you interpret this correlation, in terms of what it tells you about the pattern of contraceptive use in the sample? 

```{r}
post <- extract.samples(m13.h1)
dens( post$Rho[,1,2] )
## most have correlation around -0.7 between intercept and slope this suggest that conterception is used less in urban areas
```


## It might help to plot the mean (or median) varying effect estimates for both the intercepts and slopes, by district.  Then you can visualize the correlation and maybe more easily think through what it means to have a particular correlation.

```{r}
a1 <- apply( post$a_district , 2 , mean )
b1 <- apply( post$b_district , 2 , mean )

plot( a1 , b1 , xlab="intercept" , ylab="slope" ,
    pch=16 , col=rangi2 , ylim=c( min(b1)-0.1 , max(b1)+0.1 ) ,
    xlim=c( min(a1)-0.1 , max(a1)+0.1 ) )
points( a1 , b1 , pch=1 )
for ( i in 1:N_cafes ) 

# compute posterior mean bivariate Gaussian
Mu_est <- c( mean(post$a) , mean(post$b) )
rho_est <- mean( post$Rho[,1,2] )
sa_est <- mean( post$sigma_dist[,1] )
sb_est <- mean( post$sigma_dist[,2] )
cov_ab <- sa_est*sb_est*rho_est
Sigma_est <- matrix( c(sa_est^2,cov_ab,cov_ab,sb_est^2) , ncol=2 )
# draw contours
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
    lines(ellipse(Sigma_est,centre=Mu_est,level=l),
        col=col.alpha("black",0.2))


```

## Plotting predicted proportion of women using contraception, with urban women on one axis and rural on the other, might also help.

```{r}

nonurban <- (a1)
urban <- (a1 + b1)

plot( nonurban , urban , xlab="nonurban" , ylab="urban" ,
    pch=16 , col=rangi2 , ylim=c( min(urban)-0.1 , max(urban)+0.1 ) ,
    xlim=c( min(nonurban)-0.1 , max(nonurban)+0.1 ) )
points( nonurban , urban , pch=1 ) 

```
## julin ploting percentage in use to the correlation
## if the average effect is not zero then include as a main effect in brms 1-represents an intercept