```{r}
library(rethinking)
data(reedfrogs)
d <- reedfrogs
str(d)
```



```{r}
install.packages('ShortRead’)

```

```{r}




# make the tank cluster variable
d$tank <- 1:nrow(d)
# fit
m12.1 <- map2stan(
    alist(
        surv ~ dbinom( density , p ) ,
        logit(p) <- a_tank[tank] ,
        a_tank[tank] ~ dnorm( 0 , 5 )
), data=d )

precis(m12.1,depth=2)


m12.2 <- map2stan(
    alist(
        surv ~ dbinom( density , p ) ,
        logit(p) <- a_tank[tank] ,
        a_tank[tank] ~ dnorm( a , sigma ) ,
        a ~ dnorm(0,1) ,
        sigma ~ dcauchy(0,1)
    ), data=d , iter=4000 , chains=4 )

compare( m12.1 , m12.2 )

```

```{r}

# extract Stan samples
post <- extract.samples(m12.2)
# compute median intercept for each tank
# also transform to probability with logistic
d$propsurv.est <- logistic( apply( post$a_tank , 2 , median ) )

# display raw proportions surviving in each tank
plot( d$propsurv , ylim=c(0,1) , pch=16 , xaxt="n" ,
    xlab="tank" , ylab="proportion survival" , col=rangi2 )
axis( 1 , at=c(1,16,32,48) , labels=c(1,16,32,48) )
# overlay posterior medians
points( d$propsurv.est )
# mark posterior median probability across tanks
abline( h=logistic(median(post$a)) , lty=2 )
# draw vertical dividers between tank densities
abline( v=16.5 , lwd=0.5 )
abline( v=32.5 , lwd=0.5 )
text( 8 , 0 , "small tanks" )
text( 16+8 , 0 , "medium tanks" )
text( 32+8 , 0 , "large tanks" )

```

12M3. Re-estimate the basic Reed frog varying intercept model,but now using a Cauchy distribution in place of the Gaussian distribution for the varying intercepts.  That is,  fit this model:


```{r}
library(rethinking)
data(reedfrogs)
d <- reedfrogs

d$tank <- 1:nrow(d)


m13.2a <- map2stan(
    alist(
        surv ~ dbinom( density , p ) ,
        logit(p) <- a_tank[tank] ,
        a_tank[tank] ~ dnorm( a , sigma ) ,
        a ~ dnorm(0,1) ,
        sigma ~ dcauchy(0,1)
    ), data=d , iter=4000 , chains=4 )


m13.2b <- map2stan(
    alist(
        surv ~ dbinom( density , p ) ,
        logit(p) <- a_tank[tank] ,
        a_tank[tank] ~ cauchy( a , sigma ) ,
        a ~ dnorm(0,1) ,
        sigma ~ dcauchy(0,1)
    ), data=d , iter=4000 , chains=4 )

```

Compare the posterior means of the intercepts, αtank, to the posterior means produced in the chapter, using the customary Gaussian prior. Can you explain the pattern of di erences?

```{r}
precis(m13.2a)
precis(m13.2b)

```

12H1. In 1980, a typical Bengali woman could have 5 or more children in her lifetime. By the year 200, a typical Bengali woman had only 2 or 3. You’re going to look at a historical set of data, when contraception was widely available but many families chose not to use it.  ese data reside in data(bangladesh) and come from the 1988 Bangladesh Fertility Survey. Each row is one of 1934 women.  Tere are six variables, but you can focus on three of them for this practice problem:

```{r}
data(bangladesh) 
d <- bangladesh
head(d)

```

(1) district: ID number of administrative district each woman resided in
(2) use.contraception: An indicator (0/1) of whether the woman was using contraception
(3) urban: An indicator (0/1) of whether the woman lived in a city, as opposed to living in a rural area

 The  first thing to do is ensure that the cluster variable, district, is a contiguous set of integers. Recall that these values will be index values inside the model. If there are gaps, you’ll have parameters for which there is no data to inform them. Worse, the model probably won’t run. Look at the unique values of the district variable
 
 ```{r}
 sort(unique(d$district))
 ```
 District 54 is absent. So district isn’t yet a good index variable, because it’s not contiguous.  is is easy to  x. Just make a new variable that is contiguous.  is is enough to do it:
 
 ```{r}
  d$district_id <- as.integer(as.factor(d$district))
sort(unique(d$district_id))
 ```
 
 Now there are 60 values, contiguous integers 1 to 60.
Now, focus on predicting use.contraception, clustered by district_id. Do not include
urban just yet. Fit both (1) a traditional  xed-e ects model that uses dummy variables for district and (2) a multilevel model with varying intercepts for district. Plot the predicted proportions of women in each district using contraception, for both the  xed-e ects model and the varying-e ects model.  at is, make a plot in which district ID is on the horizontal axis and expected proportion using contraception is on the vertical. Make one plot for each model, or layer them on the same plot, as you prefer. How do the models disagree? Can you explain the pattern of disagreement? In particular, can you explain the most extreme cases of disagreement, both why they happen where they do and why the models reach di erent inferences?
 
 


